---
title: "Lab Week 8"
author: "EM August-Schmidt"
date: "March 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load the packages
```{r}
library(tidyverse) # for ggplotting
library(sf) # for reading in spatial shape file data
library(spatstat) # for point pattern analysis
library(maptools)
library(leaflet) # for making interactive (base) map plots
library(tmap) # for tmap plot
```

## Column graph of Texas oil spills 2010-2016
```{r}
oil_spills <- read_csv("oil_spills.csv")

# Subset dataframe with only data from Texas
df <- oil_spills %>%
  filter(`Accident State` == "TX" & `Accident Year` < 2017) %>%
  group_by(`Accident Year`) %>%
  summarise(Loss = sum(`Net Loss (Barrels)`))

colnames(df) <- c("Year", "Loss")

ggplot(df, aes(x=Year, y=Loss)) +
  geom_col()
```

## Leaflet plot of spill locations in Texas in 2016
```{r}
df_loc <- oil_spills %>%
  filter(`Accident State` == "TX" & `Accident Year` == 2016) %>%
  select(Latitude, Longitude, `Net Loss (Barrels)`)

colnames(df_loc) <- c("latitude","longitude", "net_loss")

oil_sf <- st_as_sf(df_loc, coords = c("longitude", "latitude"), crs = 4326) # convert data into spatial features data

# class(oil_sf) # now it's recognized as simple features data!

leaflet(oil_sf) %>%
  addTiles() %>%
  addMarkers()
```


## tmap plot with the Texas state shapefile
```{r}
states <- st_read(dsn = ".", layer = "states") # works because all files are in working directory. this reads in all files that begin 'states...'

tex_border <- states %>%
  filter(STATE_NAME == "Texas") %>%
  st_transform(4326) # give crs (coordinate reference system) for epsg value. This should match coordinate system that base map is in. 

st_crs(tex_border)

plot(tex_border)

tm_shape(tex_border) + 
  tm_polygons() +
  tm_shape(oil_sf) +
  tm_dots(size = 0.3)
```

## Convert the data to spatial points patterns (combination of point data and the bounding window)
```{r}
spill_sp <- as(oil_sf, "Spatial")

spill_ppp <- as(spill_sp, "ppp") # convert to point pattern data
class(spill_ppp)

tx_sp <- as(tex_border, "Spatial")
tx_owin <- as(tx_sp, "owin") #specifies texas border info as an Outer WINdow for point pattern analysis

all_ppp <- ppp(spill_ppp$x, spill_ppp$y, window = tx_owin)
# NB rejects any points in space that are not recognized as falling in the designated boundary window

```

## A kernel density plot for overlapping areas around plots
Beware you must choose an appropriate bandwidth of the sigma parameter (which affects the size of the grid cells) so that it is meaningful in terms of the true spatial extent of the data. 
```{r}
plot(density(all_ppp, sigma = 0.4))
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
